Convolutional Neural Nets for the MNIST Digit Dataset

### 2014-12-13 tex/paper.tex
Final edits

Though one could implement the convolutional layer directly by building the sparse network and
applying the appropriate weight updates during the back-propagation step, it is also possible to
transform this operation into the form of batch learning of the weights of a fully connected
network (discussion in section \ref{sec:full}), such that

@@

The pooling layer reduces the dimension of its input by a factor of $n$. During forward propagation, each
unit in the pooling layer looks at an $n$ x $n$ window and applies some pooling operator, either max or mean.
Though these $n$ x $n$ windows could in principle overlap, in our implementation they do not. No learning occurs
in this layer, thus back-propagation only serves to directly forward the errors to the previous layer.
If the pooling layer did max-pooling, then each pooling unit propagates the error to the input unit that
had the max value with all other errors in that window being zero. If the pooling layer does mean-pooling,
then each pooling unit propagates the value of the error divided by the number of input units to all input units.

@@

\begin{figure*}
  \includegraphics[width=\textwidth,height=8cm]{botharcs}
  \caption{The network on the left depicts the one convolutional/pooling layer architecture. The 20
  filters were of size $3 \times 3$, thus producing 20 feature maps of size $26 \times 26$. The pooling layer
  then subsamples these feature maps by a factor of 2. The network on the right depicts the two convolutional/pooling layer
  architecture. This network adds a convolutional layer with 800 filters of size $8 \times 8$. Since these filters
  are organized into kernels with 20 channels, this results in 40 $4 \times 4$ feature maps. The additional pooling
  layer then subsamples by a factor of 2. Both networks contain a subsequent fully connected layer contains 150 units,
  followed by the output layer of 10 units.}
  \label{fig:convarcs}
\end{figure*}

@@

Another way we improved the likelihood of convergence was to make each weight update of a
fully connected layer also depend on the previous weight update (this is sometimes
referred to as momentum). With this change, the weight update equation for a fully connected
layer becomes
$$\Delta W_{t} = -(\eta \frac{\partial E}{\partial W_{t}} +
\alpha \Delta W_{t-1})$$
where $\eta$ is the learning rate and $\alpha$ is the momentum.
Because many updates to $W$ have the same sign, this can effective speed up training time
\cite{WinMomentum}.
Finding good values for the momentum and learning rate seems to be
an exercise in trial-and-error, although we did find that
the learning rates we used tended to be below $0.2$.

@@

In the following we present the results of the experiments we performed. The experiments were conducted
on one of two architectures. The first architecture was composed of a single convolutional and pooling
layer followed by two fully connected layers. The convolutional layer used 20 filters of size 3x3,
while the max-pooling layer used non-overlapping windows of size 2x2. The first fully connected layer
was composed of 150 units, and the final output layer had 10 units (one per label). The second
architecture added a second set of convolutional and pooling layers before the fully connected layers.
This added convolutional layer had 800 filters of size 8x8, while the added max-pooling layer also used
on-overlapping window size of 2x2. The nonlinearity activation function across both architectures and
all layers was the hyperbolic tangent. A representation of these architectures can be seen in Figure
\ref{fig:convarcs}.

@@

The following experiments were conducted using the MNIST dataset \cite{MNIST}, which includes a training set of
60,000 28x28 handwritten digit images and a test set of 10,000 28x28 handwritten digit images.
It is important to note that both the training and test sets are composed of two sets of digits:
an easier set written by US Census Bureau employees and a harder set written by high school students.

@@

\subsubsection{Training Set of 1000 Images}

>>>>

Though one could implement the convolutional layer directly by building the sparse network and
applying the appropriate weight updates during the back-propagation step, it is also possible to
transform this operation into batch learning of the weights on a fully connected
network (discussion in section \ref{sec:full}), such that

@@

The pooling layer reduces the dimension of its input by a factor of $n$. During forward propagation, each
unit in the pooling layer looks at an $n \times n$ window and applies some pooling operator, either max or mean.
Though these $n \times n$ windows could in principle overlap, in our implementation they do not. No learning occurs
in this layer, thus back-propagation only serves to directly forward the errors to the previous layer.
If the pooling layer did max-pooling, then each pooling unit propagates the error to the input unit that
had the max value with all other errors in that window being zero. If the pooling layer does mean-pooling,
then each pooling unit propagates the value of the error divided by the number of input units to all input units.

@@

\begin{figure*}
  \includegraphics[width=\textwidth,height=8cm]{botharcs}
  \caption{The network on the left depicts the one convolutional/pooling layer architecture. The 20
  filters were of size $3 \times 3$, thus producing 20 feature maps of size $26 \times 26$. The pooling layer
  then subsamples these feature maps by a factor of 2. The network on the right depicts the two convolutional/pooling layer
  architecture. This network adds a convolutional layer with 800 filters of size $8 \times 8$. Since these filters
  are organized into kernels with 20 channels, this results in 40 $4 \times 4$ feature maps. The additional pooling
  layer then subsamples by a factor of 2. Both networks contain a subsequent fully connected layer containing 150 units,
  followed by the output layer of 10 units.}
  \label{fig:convarcs}
\end{figure*}

@@

Another way we improved the likelihood of convergence was to make each weight update of a
fully connected layer also depend on the previous weight update (this is sometimes
referred to as momentum). With this change, the weight update equation for a fully connected
layer becomes
$$\Delta W_{t} = -(\eta \frac{\partial E}{\partial W_{t}} +
\alpha \Delta W_{t-1})$$
where $\eta$ is the learning rate and $\alpha$ is the momentum.
Because many updates to $W$ have the same sign, this can effectively speed up training time
\cite{WinMomentum}.
Finding good values for the momentum and learning rate seems to be
an exercise in trial-and-error, although we did find that
the learning rates we used tended to be below $0.2$.

@@

In the following we present the results of the experiments we performed. The experiments were conducted
on one of two architectures. The first architecture was composed of a single convolutional and pooling
layer followed by two fully connected layers. The convolutional layer used 20 filters of size 3x3,
while the max-pooling layer used non-overlapping windows of size 2x2. The first fully connected layer
was composed of 150 units, and the final output layer had 10 units (one per label). The second
architecture added a second set of convolutional and pooling layers before the fully connected layers.
This added convolutional layer had 800 filters of size 8x8, while the added max-pooling layer also used
non-overlapping window size of 2x2. The nonlinearity activation function across both architectures and
all layers was the hyperbolic tangent. A representation of these architectures can be seen in Figure
\ref{fig:convarcs}.

@@

The following experiments were conducted using the MNIST dataset \cite{MNIST}, which includes a training set of
60,000 $28 \times 28$ handwritten digit images and a test set of 10,000 $28 \times 28$ handwritten digit images.
It is important to note that both the training and test sets are composed of two sets of digits:
an easier set written by US Census Bureau employees and a harder set written by high school students.

@@

\subsubsection{Training Set of 1,000 Images}
